Outliers and discrepancies

### 2024-12-08 misc/indiv.qmd
big update to results section. Added a few paragraphs to Intro.

>>>>

```{r}

s1_agg2 <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg2 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

s1_agg2 |> 
  ggplot(aes(x =log_abs_error,fill=refClass)) +
  geom_density(alpha=.5) +
  geom_boxplot(aes(x =log_abs_error,fill=refClass),y=-.1, width=.1, alpha = 0.5)

library(ggridges)
library(ggstance)
ggplot(s1_agg2, aes(y = refClass, x = log_abs_error, color = refClass)) +
  geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
                      position = position_raincloud(width = 0.05, height = 0.3,
                                                    ygap = 0.05)) +
  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
 # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
  guides(fill = "none", color = "none") +
  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "Reference Class") +
  theme(legend.position = "top")

s1_agg2 |> 
  ggplot(aes(x = log_abs_error,fill=refClass,col=refClass)) +
  ggdist::stat_halfeye(alpha=.7) +
    geom_point(
    aes(y = -.05),
    size = 4, alpha = 0.75,
    shape = 21,
    position = position_jitter(
      height = 0.01,
      seed = 2343
    )
  ) +
  facet_wrap(~ refClass,ncol=1) + ggtitle("Log-transformed error density by reference class")

s1_agg4 |> 
  ggplot(aes(x = mean_log_abs_error,fill=refClass,col=refClass)) +
  ggdist::stat_halfeye(alpha=.7) +
    geom_point(
    aes(y = -.1),
    size = 2, alpha = 0.75,
    shape = 21,
    position = position_jitter(
      height = 0.01,
      seed = 2343
    )
  ) +
  facet_wrap(~ refClass,ncol=1) + ggtitle("Log-transformed error density by reference class")

### 2024-12-08 misc/odds_ratio.qmd
big update to results section. Added a few paragraphs to Intro.

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

@@

@@

@@

>>>>

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

@@

## S1 exact match

# Assuming s1_agg is your dataframe from Experiment 1, and it's already loaded

# Fit a Bayesian logistic regression model
exact_match_model <- brm(
  matched_goal ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model.rds'))
)
summary(exact_match_model)

exact_match_model_c_add <- brm(
  matched_goal ~ refClass + calc+ (1|id) + (1|state),
  data = s1_agg,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_add.rds'))
)
summary(exact_match_model_c_add)

exact_match_model_c_int <- brm(
  matched_goal ~ refClass * calc+ (1|id) + (1|state),
  data = s1_agg,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_int.rds'))
)
summary(exact_match_model_c_int )
conditional_effects(exact_match_model_c_int)

## outliers

identify_outliers_group <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR_value <- Q3 - Q1
  lower_bound <- Q1 - 3.3 * IQR_value
  upper_bound <- Q3 + 3.3 * IQR_value
  return(x < lower_bound | x > upper_bound)
}

# 2. Apply the Outlier Detection to abs_error and log_abs_error within each group
s1_agg <- s1_agg %>%
  group_by(refClass, calc) %>%
  mutate(
    outlier_abs_error = identify_outliers_group(abs_error),
    outlier_log_abs_error = identify_outliers_group(log_abs_error)
  ) %>%
  ungroup()

#s1_agg |> relocate(outlier_abs_error,.after=abs_error) |> sort(desc(outlier_abs_error)) |> head(5)
s1_agg |> filter(outlier_abs_error) |> select(id,refClass,calc,abs_error,log_abs_error) |> # keep unique id and refClass pairs
  distinct(id,refClass)

# 3. Summarize the Number and Proportion of Outliers by Group
outlier_summary_by_group <- s1_agg %>%
  group_by(refClass, calc) %>%
  summarise(
    total_observations = n(),
    outliers_abs_error = sum(outlier_abs_error),
    percent_outliers_abs_error = round((outliers_abs_error / total_observations) * 100, 2),
    outliers_log_abs_error = sum(outlier_log_abs_error),
    percent_outliers_log_abs_error = round((outliers_log_abs_error / total_observations) * 100, 2)
  ) %>%
  ungroup()

# Display the outlier summary by group
print("Outlier Summary by Reference Class and Calculator Usage:")
print(outlier_summary_by_group)

s1_agg_no_outliers <- s1_agg %>%
  filter(!outlier_abs_error & !outlier_log_abs_error)

# Re-run the exact match comparison without outliers

# Create a contingency table of reference class vs. exact match without outliers
exact_match_table_no_outliers <- s1_agg_no_outliers %>%
  mutate(exact_match = ifelse(matched_goal, "Exact Match", "No Exact Match")) %>%
  count(refClass, exact_match) %>%
  pivot_wider(names_from = exact_match, values_from = n, values_fill = 0)

# Perform Chi-Square Test of Independence without outliers
chi_square_test_no_outliers <- chisq.test(exact_match_table_no_outliers[, -1])

# Display results
print("Chi-Square Test Results (Without Outliers):")
print(chi_square_test_no_outliers)

# Optionally, re-run logistic regression without outliers
logistic_model_no_outliers <- glm(matched_goal ~ refClass, data = s1_agg_no_outliers, family = binomial)
summary(logistic_model_no_outliers)

# Calculate Odds Ratios
odds_ratios_no_outliers <- exp(coef(logistic_model_no_outliers))
print("Odds Ratios (Without Outliers):")
print(odds_ratios_no_outliers)

# Calculate 95% Confidence Intervals for Odds Ratios without outliers
conf_intervals_no_outliers <- exp(confint(logistic_model_no_outliers))
print("95% Confidence Intervals for Odds Ratios (Without Outliers):")
print(conf_intervals_no_outliers)

s1_clean <- s1_agg %>%
  # Remove trials that are outliers in abs_error or log_abs_error
  #filter(!(outlier_abs_error | outlier_log_abs_error))
  # remove any id that has any outlier_abs_error
  filter(!(id %in% (s1_agg |> filter(outlier_abs_error) |> select(id) |> distinct() |> pull(id))))

length(unique(s1_agg$id)) 
length(unique(s1_clean$id))

# 2. Summary of Cleaned Data
cleaned_summary <- s1_clean %>%
  group_by(refClass, calc) %>%
  summarise(
    total_observations = n(),
    exact_matches = sum(matched_goal),
    percent_exact = round((exact_matches / total_observations) * 100, 2),
    mean_abs_error = mean(abs_error, na.rm = TRUE),
    mean_log_abs_error = mean(log_abs_error, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the cleaned data summary
print("Cleaned Data Summary by Reference Class and Calculator Usage:")
print(cleaned_summary)

ordinal_model_s1_clean_calc <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_clean,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
)

summary(ordinal_model_s1_clean_calc)

ordinal_model_s1_clean_calc_int <- brm(
    accuracy_level ~ refClass * calc + (1|id) + (1|state),
    data = s1_clean,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_int.rds')) 
)

summary(ordinal_model_s1_clean_calc_int)
conditional_effects(ordinal_model_s1_clean_calc_int)

# logistic regression on exact match with clean data (frequentist)
logistic_model_clean <- glm(matched_goal ~ refClass, data = s1_clean, family = binomial)
summary(logistic_model_clean)

# logistic regression on exact match with clean data (frequentist) - mixed effects
logistic_model_clean_mixed <- glmer(matched_goal ~ refClass + (1|id) + (1|state), data = s1_clean, family = binomial)
summary(logistic_model_clean_mixed)

# logistic regression on exact match with clean data (frequentist) - mixed effects
logistic_model_clean_mixed <- glmer(matched_goal ~ refClass+calc + (1|id) , data = s1_clean, family = binomial)
summary(logistic_model_clean_mixed)

# logistic regression on exact match with clean data (bayesian)
logistic_model_clean_bayes <- brm(
  matched_goal ~ refClass + (1|id) + (1|state),
  data = s1_clean,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model_clean.rds'))
)
summary(logistic_model_clean_bayes)
conditional_effects(logistic_model_clean_bayes)

## outlier from avg.

group_stats <- s1_agg |>
  group_by(refClass, calc) |>
  summarise(
    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate participant-level average abs_error
participant_stats <- s1_agg |>
  group_by(id, refClass, calc) |>
  summarise(
    avg_abs_error = mean(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Join group stats with participant stats
participant_stats <- participant_stats |>
  left_join(group_stats, by = c("refClass", "calc"))

# Identify participants to exclude based on the 2.5 SD threshold
participants_to_exclude <- participant_stats |>
  mutate(
    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
    exclude = avg_abs_error > exclusion_threshold
  ) |>
  filter(exclude)

ordinal_model_s1_clean_calc <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_clean,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
)

s1_agg_filtered <- s1_agg |>
  filter(!(id %in% participants_to_exclude$id))

length(unique(s1_agg$id)) 
length(unique(s1_agg_filtered$id))

ordinal_model_s1_clean_calc <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_agg_filtered,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
)

summary(ordinal_model_s1_clean_calc)
conditional_effects(ordinal_model_s1_clean_calc)

mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
summary(mixed_model_abs_error)
# with p value
library(lmerTest)
mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
summary(mixed_model_abs_error)

### clearn log error
```{r}

s1_clean |> 
  ggplot(aes(x = abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") +ggtitle("Error density by reference class")

# plot density of log-transformed data
s1_clean |> 
  ggplot(aes(x = log_abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") + ggtitle("Log-transformed error density by reference class")

# mixed effects regression on abs_error (frequentist)
mixed_model_abs_error <- lmer(abs_error ~ refClass + (1|id) + (1|state), data = s1_clean)
summary(mixed_model_abs_error)

mixed_model_abs_error <- lmer(abs_error ~ refClass*calc + (1|id) + (1|state), data = s1_clean)
summary(mixed_model_abs_error)

@@

# Create a new data frame with the predictor values
# new_data <- expand.grid(
#   refClass = c("kWh", "Percentage", "USD"),
#   id = unique(s2_agg1$id)[1],      # Take first id
#   state = unique(s2_agg1$state)[1], # Take first state,
#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
# )

# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
# or_table |>
#   rownames_to_column(var = "Term") |>
#   filter(!stringr::str_detect(Term, "Intercept")) |>
#   as_flextable() |>
#   set_header_labels(values = list(
#     OR = "Odds Ratio",
#     `2.5 %` = "Lower 95% CI",
#     `97.5 %` = "Upper 95% CI"
#   )) |>
#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
#   align(align = 'center', part = 'all') |>
#   fontsize(size = 10, part = "all")

@@

# Interaction s2

ordinal_model_s2_logit_int2 <- brm(
  accuracy_level ~ refClass +(rounded*pct_goal) + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int2.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int2)

conditional_effects(ordinal_model_s2_logit_int2)

ordinal_model_s2_logit_int3 <- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int3)

conditional_effects(ordinal_model_s2_logit_int3)

ordinal_model_s2_logit_c1rf_int3<- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state) + (1|calc),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1rf_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1rf_int3)

conditional_effects(ordinal_model_s2_logit_c1rf_int3)

ordinal_model_s2_logit_c1add_int3<- brm(
  accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.97), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1add_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1add_int3)

conditional_effects(ordinal_model_s2_logit_c1add_int3)

### 2024-12-08 misc/outlier.qmd
big update to results section. Added a few paragraphs to Intro.

>>>>

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

## outlier from avg.

### S1

group_stats <- s1_agg |>
  group_by(refClass, calc) |>
  summarise(
    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate participant-level average abs_error
participant_stats <- s1_agg |>
  group_by(id, refClass, calc) |>
  summarise(
    avg_abs_error = mean(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Join group stats with participant stats
participant_stats <- participant_stats |>
  left_join(group_stats, by = c("refClass", "calc"))

# Identify participants to exclude based on the 2.5 SD threshold
participants_to_exclude <- participant_stats |>
  mutate(
    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
    exclude = avg_abs_error > exclusion_threshold
  ) |>
  filter(exclude)

#saveRDS(participants_to_exclude$id,here::here("data/s1_grp_outlier_ids.rds"))

length(unique(s1_agg$id)) 
length(unique(s1_agg_filtered$id))

s1_agg_filtered <- s1_agg |>
  filter(!(id %in% participants_to_exclude$id))

summary(ordinal_model_s1_clean_calc)
conditional_effects(ordinal_model_s1_clean_calc)

mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
summary(mixed_model_abs_error)
# with p value
library(lmerTest)
mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
summary(mixed_model_abs_error)

### S2

group_stats <- s2_agg1 |>
  group_by(refClass, calc) |>
  summarise(
    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate participant-level average abs_error
participant_stats <- s2_agg1 |>
  group_by(id, refClass, calc) |>
  summarise(
    avg_abs_error = mean(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Join group stats with participant stats
participant_stats <- participant_stats |>
  left_join(group_stats, by = c("refClass", "calc"))

# Identify participants to exclude based on the 2.5 SD threshold
participants_to_exclude <- participant_stats |>
  mutate(
    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
    exclude = avg_abs_error > exclusion_threshold
  ) |>
  filter(exclude)

s2_agg_filtered <- s2_agg1 |>
  filter(!(id %in% participants_to_exclude$id))

#saveRDS(participants_to_exclude$id,here::here("data/s2_grp_outlier_ids.rds"))

mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s2_agg_filtered)
summary(mixed_model_abs_error)
# with p value
library(lmerTest)
mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s2_agg_filtered)
summary(mixed_model_abs_error)

library(lmerTest)
mixed_model_abs_error <- lmer(log_abs_error ~ refClass*rounded*pct_goal + calc+ (1|id) + (1|state), data = s2_agg_filtered)
summary(mixed_model_abs_error)

ordinal_model_s2_logit <- brm(
  accuracy_level ~ (refClass*rounded*pct_goal)+calc + (1|id)+ (1|state),
  data = s2_agg_filtered,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_grpclean_calc_int.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit)

conditional_effects(ordinal_model_s2_logit)

# old
